Learning growth mechanisms of tail realistic preferential attachment models from network degree distributions

### 2025-10-03 Files/fixed_paper.tex
Final changes for preprint

\raggedbottom
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{\textbf{Fig.}}
\else
  \newcommand\figurename{\textbf{Fig.}}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\usepackage{amsthm}
\theoremstyle{plain}
\theoremstyle{plain}
\theoremstyle{remark}
\AtBeginDocument{\renewcommand*{\proofname}{Proof}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Learning growth mechanisms of tail realistic preferential attachment models from network degree distributions},
  pdfauthor={Thomas William Boughen; Clement Lee; Vianey Palacios Ramírez},
  pdfkeywords={Extremes · Discrete extremes · Power law · Preferential
attachment},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

% keywords
\keywords{Extremes · Discrete extremes · Power law · Preferential
attachment}

@@

Consider a continuous time branching process \(\zeta(t)\) driven by a
Markovian pure birth process, with \(\zeta(0)=0\) and birth rates
depending on a non-negative function \(b(\cdot)\):

\[
\Pr(\zeta(t+\text{d}t)=k+1|\zeta(t)=k) = b(k)\text{d}t + o(\text{d}t).
\]

Denote the density of the point process associated with the pure birth
process corresponding to the growth of an individual node by
\(\rho(t)\), and its Laplace transform by
\(\hat \rho(\lambda)\coloneq \int_0^\infty e^{-\lambda t}\rho(t)\text{d}t\).
We denote the Malthusian parameter of this process by \(\lambda^*\),
that is \(\lambda^*\) satisfies \(\hat\rho(\lambda^*) = 1\).

@@

Here, \(\lambda^*\) is the Malthusian parameter of the corresponding
branching process. See Appendix \ref{sec-proofs} for the details of the
proof.

@@

\begin{refremark}
If \(\lim_{k\rightarrow\infty}b(k)<\infty\), then \(\Omega(F, k) = 0\).
Furthermore, \(F\) is recovered to the Gumbel MDA but does not belong to
the Gumbel MDA itself as it is not long-tailed. See Appendix
\ref{sec-proofs} for the proof.

@@

Corollary~\ref{cor-omega2} is a direct consequence of
Proposition~\ref{prp-omega} and the results of \citet{shimura12}, and it
aligns with previous findings in the literature. In particular it is
consistent with \citet{krapivsky01}, who showed that a sub-linear
preference function leads to a light-tailed distributions, since for
\(b(k)=k^\alpha\) with \(\alpha<1\), we have
\(\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=0\). Conversely, it also agrees
with the limiting behaviour of the BA model, which produces a regularly
degree distribution with tail index 2; for the preference function
\(b(k)=k+1\), \(\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=1\), giving a
tail index of \(\lambda^*=2\).

@@

\begin{equation}\phantomsection\label{eq-rho}{
\hat\rho(\lambda) = \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{i^\alpha + \varepsilon}{\lambda+i^\alpha + \varepsilon} + \left(\frac{k_0^\alpha + \varepsilon}{\lambda-\beta}\right)\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda + i^\alpha + \varepsilon}, 
}\end{equation} which has to be solved numerically for most parameter
choices. Also, note that \(\lambda>\beta\). See Appendix
\ref{sec-proofs} for the derivations of Equation~\ref{eq-rho}.

@@

In order to carry out the Bayesian inference of the parameters we must
first assign priors, for the parameters defined on the positive real
line (\(\alpha, \beta,\varepsilon\)) we use diffuse gamma priors and for
\(k_0\) we use a discrete uniform prior over a large and reasonable
range of values.

@@

The procedure for recovering the parameters begins with simulating a
network from the model with \(N=100,000\) vertices and \(m=1\) given
some set of parameters
\(\pmb\theta = (\alpha, \beta, \varepsilon, k_0)\), obtaining the degree
counts and using the likelihood from the previous section alongside the
priors:

@@

@@

\subsection{\texorpdfstring{Proof of
Remark~\ref{rem-omega}}{Proof of Remark~}}\label{proof-of-rem-omega}

As it is assumed throughout that \(b(k)\) is non-decreasing, we consider
the two cases where \(b(k)\) is bounded and unbounded. When \(b(k)\) is
bounded, by the monotone convergence theorem,
\(\lim_{k\rightarrow\infty}b(k) = c\) for some \(c>0\), which implies
that \(\lim_{k\rightarrow\infty}\Omega(F,k)= 0\). This proves the first
part of Remark~\ref{rem-omega}. Now, using
Equation~\ref{eq-surv-origin},
\(\frac{\bar{F}(k+1)}{\bar{F}(k)}=\frac{b(k+1)}{\lambda^{*}+b(k+1)}\rightarrow\frac{c}{\lambda^{*}+c}<1\)
as \(k\rightarrow\infty\). This means that \(F\) is not long-tailed, and
therefore is only recovered to the Gumbel MDA.

\subsection{\texorpdfstring{Proof of
Proposition~\ref{prp-omega}}{Proof of Proposition~}}\label{proof-of-prp-omega}

@@

\subsection{\texorpdfstring{Derivation of
Equation~\ref{eq-rho}}{Derivation of Equation~}}\label{derivation-of-eq-rho}

>>>>

\raggedbottom
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{\textbf{Fig.}}
\else
  \newcommand\figurename{\textbf{Fig.}}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\usepackage{amsthm}
\theoremstyle{plain}
\theoremstyle{plain}
\theoremstyle{remark}
\AtBeginDocument{\renewcommand*{\proofname}{Proof}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Learning growth mechanisms of tail realistic preferential attachment models from network degree distributions},
  pdfauthor={Thomas William Boughen; Clement Lee; Vianey Palacios Ramírez},
  pdfkeywords={Network degrees · Discrete extremes · Power law ·
Preferential attachment},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

% keywords
\keywords{Network degrees · Discrete extremes · Power law · Preferential
attachment}

@@

Consider a continuous time branching process \(\zeta(t)\) driven by a
Markovian pure birth process, with \(\zeta(0)=0\) and birth rates
depending on a non-negative function \(b(\cdot)\): \[
\Pr(\zeta(t+\text{d}t)=k+1|\zeta(t)=k) = b(k)\text{d}t + o(\text{d}t).
\] Denote the density of the point process associated with the pure
birth process corresponding to the growth of an individual node by
\(\rho(t)\), and its Laplace transform by
\(\hat \rho(\lambda)\coloneq \int_0^\infty e^{-\lambda t}\rho(t)\text{d}t\).
We denote the Malthusian parameter of this process by \(\lambda^*\),
that is \(\lambda^*\) satisfies \(\hat\rho(\lambda^*) = 1\).

@@

Here, \(\lambda^*\) is the Malthusian parameter of the corresponding
branching process. See Appendix \ref{sec-prpproof} for the details of
the proof.

@@

\begin{refremark}
If \(\lim_{k\rightarrow\infty}b(k)<\infty\), then \(\Omega(F, k) = 0\).
Furthermore, \(F\) is recovered to the Gumbel MDA but does not belong to
the Gumbel MDA itself as it is not long-tailed. See Appendix
\ref{sec-remproof} for the proof.

@@

Corollary~\ref{cor-omega2} is a direct consequence of
Proposition~\ref{prp-omega} and the results of \citet{shimura12}, and it
aligns with previous findings in the literature. In particular it is
consistent with \citet{krapivsky01}, who showed that a sub-linear
preference function leads to a light-tailed distributions, since for
\(b(k)=k^\alpha\) with \(\alpha<1\), we have
\(\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=0\). Conversely, it also agrees
with the limiting behaviour of the BA model, which produces a regularly
degree distribution with tail index 2; for the preference function
\(b(k)=k+1\), \(\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=1\), giving a
tail index of \(\lambda^*=2\). We can also obtain the afforementioned
results for UA, taking \(b(k)=c\) it is clear that
\(\lim_{k\rightarrow\infty}b(k)<\infty\) and therefore by
Remark~\ref{rem-omega} we have tha
\(\lim_{k\rightarrow\infty}\Omega(F,k)=0\) meaning \(F\)'s tail is not
regularly varying and is in the Gumbel domain as \(b(k)\) is also
long-tailed.

@@

\begin{equation}\phantomsection\label{eq-rho}{
\hat\rho(\lambda) = \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{i^\alpha + \varepsilon}{\lambda+i^\alpha + \varepsilon} + \left(\frac{k_0^\alpha + \varepsilon}{\lambda-\beta}\right)\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda + i^\alpha + \varepsilon}, 
}\end{equation} which has to be solved numerically for most parameter
choices. Also, note that \(\lambda>\beta\). See Appendix
\ref{sec-rhoproof} for the derivations of Equation~\ref{eq-rho}.

@@

The methodology for the Bayesian inference of the parameters is detailed
in Section~\ref{sec-sim}.

@@

The procedure for recovering the parameters begins with simulating a
network from the model with \(N=100,000\) vertices and \(m=1\) given
some set of parameters
\(\pmb\theta = (\alpha, \beta, \varepsilon, k_0)\). In order to carry
out the Bayesian inference of the parameters we must first assign
priors, for the parameters defined on the positive real line
(\(\alpha, \beta,\varepsilon\)) we use diffuse gamma priors and for
\(k_0\) we use a discrete uniform prior over a large and reasonable
range of values, specifically we use the priors below:

@@

The parameter inferences that follow were obtained using the same priors
and MCMC sampling method as detailed in Section~\ref{sec-sim}.

@@

\subsection{\texorpdfstring{Proof of
Remark~\ref{rem-omega}}{Proof of Remark~}}\label{sec-remproof}

As it is assumed throughout that \(b(k)\) is non-decreasing, we consider
the two cases where \(b(k)\) is bounded and unbounded. When \(b(k)\) is
bounded, by the monotone convergence theorem,
\(\lim_{k\rightarrow\infty}b(k) = c\) for some \(c>0\), which implies
that
\(\lim_{k\rightarrow\infty}\Omega(F,k)=\left(\log\left[1+\frac{\lambda^{*}}{c}\right]\right)^{-1}-\left(\log\left[1+\frac{\lambda^{*}}{c}\right]\right)^{-1}= 0\).
This proves the first part of Remark~\ref{rem-omega}. Now, using
Equation~\ref{eq-surv-origin},
\(\frac{\bar{F}(k+1)}{\bar{F}(k)}=\frac{b(k+1)}{\lambda^{*}+b(k+1)}\rightarrow\frac{c}{\lambda^{*}+c}<1\)
as \(k\rightarrow\infty\). This means that \(F\) is not long-tailed, and
therefore is only recovered to the Gumbel MDA.

\subsection{\texorpdfstring{Proof of
Proposition~\ref{prp-omega}}{Proof of Proposition~}}\label{sec-prpproof}

@@

\subsection{\texorpdfstring{Derivation of
Equation~\ref{eq-rho}}{Derivation of Equation~}}\label{sec-rhoproof}

### 2025-10-03 Files/paper.qmd
Final changes for preprint

---
title: "Learning growth mechanisms of tail realistic preferential attachment models from network degree distributions"
crossref:
  fig-title: '**Fig.**'        
format:
  nature-pdf:
    fontsize: 10pt
    indent: true
    # journal.cite-style is included in the tex file but ignored by pandoc if 
    # cite-method is not `natbib`.
    journal:
      cite-style: sn-basic
    # `citeproc` is the pandoc default. Set `cite-method: natbib` if required
    # to use the bst styles from the upstream template.
    cite-method: natbib
    keep-tex: true
    number-sections: true
author:
  - name: Thomas William Boughen
    orcid: 0009-0005-1979-2479
    email: t.w.boughen1@newcastle.ac.uk
    corresponding: true
    affiliations:
      - name: Newcastle University
        id: 1
        department: School of Mathematics, Statistics and Physics
  - name: Clement Lee
    affiliations:
      - ref: 1
  - name: Vianey Palacios Ramírez
    affiliations:
      - ref: 1
abstract: |
  Identifying the generating mechanism of a network is challenging as, more often than not, only snapshots are available, but not the full evolution. One candidate for the generating mechanism is preferential attachment which, in its simplest form, results in a degree distribution that follows the power law. Consequently, the growth of real-life networks that display such power-law behaviour is commonly modelled by preferential attachment. The ubiquity of the power law has been challenged by the presence of alternatives with comparable performance, as well as the recent findings that the tail of the degree distribution is often lighter than implied by the body, whilst still being regularly varying. In this paper, we propose a preferential attachment model with a flexible preference function. Using methods for discrete extremes, we characterise the tail behaviour of the limiting degree distribution directly by the form of the preference function. Directly relating the tail index to the model parameters enables them to be inferred when fitting to degree distributions alone, which is supported by simulation studies. Results of applications to real data are promising and comparable to alternatives.
keywords: Extremes · Discrete extremes · Power law · Preferential attachment
msc: 
 - 05C80
 - 60J85
 - 62G32
bibliography: refs.bib  
keep-tex: true
pdf-engine: xelatex
---

@@

Consider a continuous time branching process $\zeta(t)$ driven by a Markovian pure birth process, with $\zeta(0)=0$ and birth rates depending on a non-negative function $b(\cdot)$:

$$
\Pr(\zeta(t+\text{d}t)=k+1|\zeta(t)=k) = b(k)\text{d}t + o(\text{d}t).
$$

Denote the density of the point process associated with the pure birth process corresponding to the growth of an individual node by $\rho(t)$, and its Laplace transform by $\hat \rho(\lambda)\coloneq \int_0^\infty e^{-\lambda t}\rho(t)\text{d}t$. We denote the Malthusian parameter of this process by $\lambda^*$, that is $\lambda^*$ satisfies $\hat\rho(\lambda^*) = 1$.

@@

Here, $\lambda^*$ is the Malthusian parameter of the corresponding branching process. See Appendix [-@sec-proofs] for the details of the proof.
:::

::: {#rem-omega}
If $\lim_{k\rightarrow\infty}b(k)<\infty$, then $\Omega(F, k) = 0$. Furthermore, $F$ is recovered to the Gumbel MDA but does not belong to the Gumbel MDA itself as it is not long-tailed. See Appendix [-@sec-proofs] for the proof.
:::

@@

@cor-omega2 is a direct consequence of @prp-omega and the results of @shimura12, and it aligns with previous findings in the literature. In particular it is consistent with @krapivsky01, who showed that a sub-linear preference function leads to a light-tailed distributions, since for $b(k)=k^\alpha$ with $\alpha<1$, we have $\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=0$. Conversely, it also agrees with the limiting behaviour of the BA model, which produces a regularly degree distribution with tail index 2; for the preference function $b(k)=k+1$, $\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=1$, giving a tail index of $\lambda^*=2$.

@@

```{r, echo=FALSE, out.width="119mm", warning=FALSE, message=FALSE, cache=TRUE}
#| fig-width: 6
#| fig-asp: 0.7
#| fig-cap: Example construction of the preference function in @eq-pref
#| label: fig-ex
library(ggbrace)
x = seq(0,10,l=100)
k0 = 5
a = 0.5
eps = 0.5
y2 = ifelse(x<=k0, (eps + x^1) , (x-k0) + k0^1 + eps) +  (k0^a - k0)
y3 = ifelse(x<=k0, (eps + x^a), (x-k0) + k0^a + eps)
y4 = eps + x^a

@@

$$
\hat\rho(\lambda) = \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{i^\alpha + \varepsilon}{\lambda+i^\alpha + \varepsilon} + \left(\frac{k_0^\alpha + \varepsilon}{\lambda-\beta}\right)\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda + i^\alpha + \varepsilon}, 
$$ {#eq-rho}
which has to be solved numerically for most parameter choices. Also, note that $\lambda>\beta$. See Appendix [-@sec-proofs] for the derivations of @eq-rho.

@@

In order to carry out the Bayesian inference of the parameters we must first assign priors, for the parameters defined on the positive real line ($\alpha, \beta,\varepsilon$) we use diffuse gamma priors and for $k_0$ we use a discrete uniform prior over a large and reasonable range of values.

@@

The procedure for recovering the parameters begins with simulating a network from the model with $N=100,000$ vertices and $m=1$ given some set of parameters $\pmb\theta = (\alpha, \beta, \varepsilon, k_0)$, obtaining the degree counts and using the likelihood from the previous section alongside the priors:

@@

@@

## Proof of @rem-omega

As it is assumed throughout that $b(k)$ is non-decreasing, we consider the two cases where $b(k)$ is bounded and unbounded. When $b(k)$ is bounded, by the monotone convergence theorem, $\lim_{k\rightarrow\infty}b(k) = c$ for some $c>0$, which implies that $\lim_{k\rightarrow\infty}\Omega(F,k)= 0$. This proves the first part of @rem-omega. Now, using @eq-surv-origin,  $\frac{\bar{F}(k+1)}{\bar{F}(k)}=\frac{b(k+1)}{\lambda^{*}+b(k+1)}\rightarrow\frac{c}{\lambda^{*}+c}<1$ as $k\rightarrow\infty$. This means that $F$ is not long-tailed, and therefore is only recovered to the Gumbel MDA.

## Proof of @prp-omega

@@

## Derivation of @eq-rho

>>>>

---
title: "Learning growth mechanisms of tail realistic preferential attachment models from network degree distributions"
crossref:
  fig-title: '**Fig.**'        
format:
  nature-pdf:
    fontsize: 10pt
    indent: true
    # journal.cite-style is included in the tex file but ignored by pandoc if 
    # cite-method is not `natbib`.
    journal:
      cite-style: sn-basic
    # `citeproc` is the pandoc default. Set `cite-method: natbib` if required
    # to use the bst styles from the upstream template.
    cite-method: natbib
    keep-tex: true
    number-sections: true
author:
  - name: Thomas William Boughen
    orcid: 0009-0005-1979-2479
    email: t.w.boughen1@newcastle.ac.uk
    corresponding: true
    affiliations:
      - name: Newcastle University
        id: 1
        department: School of Mathematics, Statistics and Physics
  - name: Clement Lee
    affiliations:
      - ref: 1
  - name: Vianey Palacios Ramírez
    affiliations:
      - ref: 1
abstract: |
  Identifying the generating mechanism of a network is challenging as, more often than not, only snapshots are available, but not the full evolution. One candidate for the generating mechanism is preferential attachment which, in its simplest form, results in a degree distribution that follows the power law. Consequently, the growth of real-life networks that display such power-law behaviour is commonly modelled by preferential attachment. The ubiquity of the power law has been challenged by the presence of alternatives with comparable performance, as well as the recent findings that the tail of the degree distribution is often lighter than implied by the body, whilst still being regularly varying. In this paper, we propose a preferential attachment model with a flexible preference function. Using methods for discrete extremes, we characterise the tail behaviour of the limiting degree distribution directly by the form of the preference function. Directly relating the tail index to the model parameters enables them to be inferred when fitting to degree distributions alone, which is supported by simulation studies. Results of applications to real data are promising and comparable to alternatives.
keywords: Network degrees · Discrete extremes · Power law · Preferential attachment
msc: 
 - 05C80
 - 60J85
 - 62G32
bibliography: refs.bib  
keep-tex: true
pdf-engine: xelatex
---

@@

Consider a continuous time branching process $\zeta(t)$ driven by a Markovian pure birth process, with $\zeta(0)=0$ and birth rates depending on a non-negative function $b(\cdot)$:
$$
\Pr(\zeta(t+\text{d}t)=k+1|\zeta(t)=k) = b(k)\text{d}t + o(\text{d}t).
$$
Denote the density of the point process associated with the pure birth process corresponding to the growth of an individual node by $\rho(t)$, and its Laplace transform by $\hat \rho(\lambda)\coloneq \int_0^\infty e^{-\lambda t}\rho(t)\text{d}t$. We denote the Malthusian parameter of this process by $\lambda^*$, that is $\lambda^*$ satisfies $\hat\rho(\lambda^*) = 1$.

@@

Here, $\lambda^*$ is the Malthusian parameter of the corresponding branching process. See Appendix [-@sec-prpproof] for the details of the proof.
:::

::: {#rem-omega}
If $\lim_{k\rightarrow\infty}b(k)<\infty$, then $\Omega(F, k) = 0$. Furthermore, $F$ is recovered to the Gumbel MDA but does not belong to the Gumbel MDA itself as it is not long-tailed. See Appendix [-@sec-remproof] for the proof.
:::

@@

@cor-omega2 is a direct consequence of @prp-omega and the results of @shimura12, and it aligns with previous findings in the literature. In particular it is consistent with @krapivsky01, who showed that a sub-linear preference function leads to a light-tailed distributions, since for $b(k)=k^\alpha$ with $\alpha<1$, we have $\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=0$. Conversely, it also agrees with the limiting behaviour of the BA model, which produces a regularly degree distribution with tail index 2; for the preference function $b(k)=k+1$, $\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=1$, giving a tail index of $\lambda^*=2$. We can also obtain the afforementioned results for UA, taking $b(k)=c$ it is clear that $\lim_{k\rightarrow\infty}b(k)<\infty$ and therefore by @rem-omega we have tha $\lim_{k\rightarrow\infty}\Omega(F,k)=0$ meaning $F$'s tail is not regularly varying and is in the Gumbel domain as $b(k)$ is also long-tailed.

@@

```{r, echo=FALSE, out.width="119mm", warning=FALSE, message=FALSE, cache=TRUE}
#| fig-width: 6
#| fig-asp: 0.55
#| fig-cap: Example construction of the preference function in @eq-pref
#| label: fig-ex
library(ggbrace)
x = seq(0,10,l=100)
k0 = 5
a = 0.5
eps = 0.5
y2 = ifelse(x<=k0, (eps + x^1) , (x-k0) + k0^1 + eps) +  (k0^a - k0)
y3 = ifelse(x<=k0, (eps + x^a), (x-k0) + k0^a + eps)
y4 = eps + x^a

@@

$$
\hat\rho(\lambda) = \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{i^\alpha + \varepsilon}{\lambda+i^\alpha + \varepsilon} + \left(\frac{k_0^\alpha + \varepsilon}{\lambda-\beta}\right)\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda + i^\alpha + \varepsilon}, 
$$ {#eq-rho}
which has to be solved numerically for most parameter choices. Also, note that $\lambda>\beta$. See Appendix [-@sec-rhoproof] for the derivations of @eq-rho.

@@

The methodology for the Bayesian inference of the parameters is detailed in @sec-sim.

@@

The procedure for recovering the parameters begins with simulating a network from the model with $N=100,000$ vertices and $m=1$ given some set of parameters $\pmb\theta = (\alpha, \beta, \varepsilon, k_0)$. In order to carry out the Bayesian inference of the parameters we must first assign priors, for the parameters defined on the positive real line ($\alpha, \beta,\varepsilon$) we use diffuse gamma priors and for $k_0$ we use a discrete uniform prior over a large and reasonable range of values, specifically we use the priors below:

@@

The parameter inferences that follow were obtained using the same priors and MCMC sampling method as detailed in @sec-sim.

@@

## Proof of @rem-omega {#sec-remproof}

As it is assumed throughout that $b(k)$ is non-decreasing, we consider the two cases where $b(k)$ is bounded and unbounded. When $b(k)$ is bounded, by the monotone convergence theorem, $\lim_{k\rightarrow\infty}b(k) = c$ for some $c>0$, which implies that $\lim_{k\rightarrow\infty}\Omega(F,k)=\left(\log\left[1+\frac{\lambda^{*}}{c}\right]\right)^{-1}-\left(\log\left[1+\frac{\lambda^{*}}{c}\right]\right)^{-1}= 0$. This proves the first part of @rem-omega. Now, using @eq-surv-origin,  $\frac{\bar{F}(k+1)}{\bar{F}(k)}=\frac{b(k+1)}{\lambda^{*}+b(k+1)}\rightarrow\frac{c}{\lambda^{*}+c}<1$ as $k\rightarrow\infty$. This means that $F$ is not long-tailed, and therefore is only recovered to the Gumbel MDA.

## Proof of @prp-omega {#sec-prpproof}

@@

## Derivation of @eq-rho {#sec-rhoproof}

### 2025-10-03 Files/paper.tex
Final changes for preprint

\raggedbottom
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{\textbf{Fig.}}
\else
  \newcommand\figurename{\textbf{Fig.}}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\usepackage{amsthm}
\theoremstyle{plain}
\newtheorem{corollary}{Corollary}[section]
\theoremstyle{plain}
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{remark}
\AtBeginDocument{\renewcommand*{\proofname}{Proof}}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\newtheorem{refremark}{Remark}[section]
\newtheorem{refsolution}{Solution}[section]
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Learning growth mechanisms of tail realistic preferential attachment models from network degree distributions},
  pdfauthor={Thomas William Boughen; Clement Lee; Vianey Palacios Ramírez},
  pdfkeywords={Extremes · Discrete extremes · Power law · Preferential
attachment},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

>>>>

\raggedbottom
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{\textbf{Fig.}}
\else
  \newcommand\figurename{\textbf{Fig.}}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\usepackage{amsthm}
\theoremstyle{plain}
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{plain}
\newtheorem{corollary}{Corollary}[section]
\theoremstyle{remark}
\AtBeginDocument{\renewcommand*{\proofname}{Proof}}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\newtheorem{refremark}{Remark}[section]
\newtheorem{refsolution}{Solution}[section]
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Learning growth mechanisms of tail realistic preferential attachment models from network degree distributions},
  pdfauthor={Thomas William Boughen; Clement Lee; Vianey Palacios Ramírez},
  pdfkeywords={Network degrees · Discrete extremes · Power law ·
Preferential attachment},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}
